# Inflammation Index R Package
---

## Setup

---

### Step 1: Installation

In R the InflammationIndex package can be installed first by
installing and loading devtools to enable installing packages directly from GitHub, 
then installing the InflammationIndex package from GitHub

```{r, eval = False}
require(devtools)
install_github("BrainEnergyLab/Inflammation-Index")
```

Then load in the package.

```{r}
require(InflammationIndex)
```

---

## Use with the Microglia Morphology Analysis Fiji Plugin

---

### Step 2: Loading in Data (morphPreProcessing() function)

After users have run the Microglia Morphology Analysis Fiji plugin,
the morphPreProcessing() function can be used to read in and collate all the morphological
metrics extracted.This function requires users to specify:
- pixelSize: The pixel size in microns (assuming a square pixel)
- morphologyWD: The file path to the 'Output' folder of the 'working directory' of the Fiji plugin passed as a string 
- animalIDs: A string vector where each element is the ID of an animal that users want to extract data for
- treatmentIDs: A string vector where each element is the ID of the treatment applied that users want to extract data for

Optional inputs:
- TCSExclude: a numeric vector of mask size values that users want to exclude from the data collation
  - Defaults to NULL (no values are excluded)
- useFrac: a boolean that indicates whether the user wants to include data from FracLac in their collation
  - Defaults to False

The output of the function is a data.table where each row is a cell and the columns indicate
morphological metrics.

```{r}
pixelSize = 0.58 # Pixel size in microns
morphologyWD = "/Microglial Morphology/Output" # Output directory of the MicroMorph.ijm script as a string
animalIDs = c('HIPP5', 'HIPP6', 'HIPP7') # Vector of strings identifying the names of the animals images were captured from and matching the names of the Animal level folders
treatmentIDs = c('Pre-LPS', 'Post-LPS') # Vector of strings identifying different treatments / timepoints and matching the names of the Treatment level folders
useFrac = T # Boolean indicating whether to use the output of the FracLac plugin
TCSExclude = NULL # String vector of mask sizes to exclude from the preprocessing function, can also take NULL
```

#### Retrieving animal and treatment IDs from the Fiji plugin folder structure

To avoid having to manually input the animal and treatment IDs, users can use the
getAnimalAndTreatmentIDs() function to return a list, the only input is:
- imageStorageDirectory: the file path to the 'image storage directory' that users used with the
Fiji plugin, passed as a string

The function returns a list with two elements:
- treatmentIDs: a string vector of treatment IDs
- animalIDs: a string vector of animal IDs
These can be passed directly to the treatmentIDs and animalIDs arguments in morphPreProcessing().

```{r include = False}

imageStorageDirectory = '/Users/devin.clarke/Google Drive/Microglia Morphology Analysis Plugin - ImageJ Example Directory and Input Data/Image Storage Directory'
morphologyWD = '/Users/devin.clarke/Google Drive/Microglia Morphology Analysis Plugin - ImageJ Example Directory and Input Data/Working Directory/Output'

imageStorageDirectoryOld = '/Users/devin.clarke/Dropbox (Brain Energy Lab)/Everything/2P data/Devin/'
morphologyWDOld = '/Users/devin.clarke/Dropbox (Brain Energy Lab)/Everything/Devin/2P Data Analysis/Microglial Morphology/Output/'

exampleData = '/Users/devin.clarke/Google Drive/Inflammation Index R Package Example Data/example_morphpreprocessing_output.csv'

```

Here we're using the 'Image Storage Directory' in the example data found at our [Fji example data directory](https://drive.google.com/drive/folders/1axi8uFPbbyMObYko62jP14De9DCrQeUT?usp=sharing)

```{r}

idList = getAnimalAndTreatmentIDs(imageStorageDirectory)
treatmentIDs = idList$treatmentIDs
animalIDs = idList$animalIDs

idList

```

Here we're using the 'Working Directory' in the example data found at our [Fji example data directory](https://drive.google.com/drive/folders/1axi8uFPbbyMObYko62jP14De9DCrQeUT?usp=sharing)

```{r}
output = 
  morphPreProcessing(
    pixelSize = pixelSize, morphologyWD = morphologyWD, 
    animalIDs = animalIDs, treatmentIDs = treatmentIDs,
    useFrac = useFrac)
```

The colums in the output table:
- Animal: animal ID 
- Treatment: treatment ID
- TCSValue:  The mask size value this row's metrics relate to
- UniqueID is a unique identifier for each row (a combination of animal, treatment, mask size, and name of the mask file)
- CellNo (second to last column) is an identifier for each cell that is duplicated if a cell
has measurements taken at multiple TCSValue levels. 

All other columns are morphological metrics.

```{r}
head(output)
```

### Step 3: Constructing the Inflammation Index (constructInfInd() function)

Once users have an output from the morphPreProcessing() function, they can use this
to generate an inflammation index based on training conditions using the constructInfInd() function.
This function takes three required arguments:
- inDat: this is the data.table output by the morphPreProcessing() function
- LPSGroups: this is a string vector of the treatmentIDs use to identify the training conditions
  - The length of this vector should be 2
- method: this is a string identifying which method users want to use to optimise the Inflammation Index
  -'p value' uses the smallest p value
  -'AUC' uses a ROC-AUC analysis
  
In addition there is one optional argument:
- noDesc: this is an integer vector of the number of 'best' descriptors to compare to one another
  - Defaults to 1:15
  
The function returns a single PCA object.

```{r}
inDat = output # The output of the morphPreProcessing() function
LPSGroups = c('Pre-LPS', 'Post-LPS') # Vector of strings of the positive control conditions
method = 'p value' # The method to use to refine the inflammation index, can also take the value 'AUC
noDesc = 1:15 # A vector of integers where we compare inflammation indices made up of the N best descriptors and compare them
  # E.g. when noDesc = c(1,2,3) we compare the inflammation indices made up of the 1 best discriminators, to one made of the 2 best discriminators, to one made of the 3 best discrminators.
```

Here we're going to run this function on the example morphPreProccessing output table
found [here](https://drive.google.com/file/d/1dtgZZuBTPg-uJaWxZy8bOav8mSJs-msY/view?usp=sharing). 
It's worth noting that this data was collated with a legacy version of the morphPreProcessing function
and so has slightly different column names and ID column values.

```{r include = False}

inDat = fread(exampleData)

```

```{r}

head(inDat)

```
The constructInfInd() function will print out the mask size, and number of descriptors,
that created the Inflammation Index that was most sensitive to the differences in morphology
between the LPSGroups conditions according to the method passed in the method argument. It will
print the value from the method argument.

```{r}
infIndOut = 
  constructInfInd(inDat = inDat,
                  LPSGroups = c('D56', 'LPS'),
                  method = 'p value',
                  noDesc = 5:15)
```

### Step 4: Apply the Inflammation Index to test data (apply_inf_ind())

Using the mask size printed by the constructInfInd, users should run the
MicroMorph.ijm script on the dataset they want to apply the constructed
Inflammation Index to. In the ‘Generate masks for cells’ step users can
accomplish this by setting the upper and lower limits of the mask sizes
to the optimal mask size.

## Step 3: Running InflammationIndex on Experimental Data

Once the MicroMorph.ijm output has been generated for the experimental
data, users must run the morphPreProcessing function on this
data

``` r
non_positive_treatments = c('EXP_CONDITION_1', 'EXP_CONDITION_2')
```

``` r
# Run the morphPreProcessing() function this time on experimental data - as specified in the treatmentIDs argument in this case
output_exp_data = 
  morphPreProcessing(
    pixelSize = pixelSize, morphologyWD = morphologyWD, 
    animalIDs = animalIDs, treatmentIDs = non_positive_treatments,
    useFrac = useFrac)
```

Following this, users use the apply\_inf\_ind() function to apply to
inflammation index generated earlier. This outputs a data table that is
the same as the output\_exp\_data data table but with an extra column
added that is the inflammation index calculated for each cell. This
inflammation index metric can then be used to analyse whether the
experimental conditions have an effect on microglial morphology.

``` r
data_with_inf_index = apply_inf_ind(infIndOut, output_exp_data)
head(data_with_inf_index)
```

    ##    TCS Animal Treatment                                                UniqueID
    ## 1: 500 HIPP17       D30     HIPP17D30CANDIDATEMASKFORSUBSTACK(21-30)X100Y249500
    ## 2: 500   CE1L   D2HOURS   CE1LD2HOURSCANDIDATEMASKFORSUBSTACK(21-30)X100Y348500
    ## 3: 500   BU2L        D1         BU2LD1CANDIDATEMASKFORSUBSTACK(21-30)X100Y61500
    ## 4: 500   BG1L        D7        BG1LD7CANDIDATEMASKFORSUBSTACK(21-30)X101Y116500
    ## 5: 500 HIPP12   D4HOURS HIPP12D4HOURSCANDIDATEMASKFORSUBSTACK(21-30)X102Y134500
    ## 6: 500   BR1R       D-1       BR1RD-1CANDIDATEMASKFORSUBSTACK(21-30)X102Y192500
    ##    CellParametersPerimeter CellSpread Eccentricity Roundness SomaSize MaskSize
    ## 1:                 327.259     24.864        1.758     0.053   51.133  450.776
    ## 2:                 263.188     21.937        1.632     0.067   37.004  372.058
    ## 3:                 381.371     24.091        1.233     0.049   66.271  562.461
    ## 4:                 196.199     15.475        1.184     0.097   54.160  298.050
    ## 5:                 178.576     16.658        1.562     0.123   25.903  312.516
    ## 6:                 270.910     19.957        1.362     0.062   45.750  359.612
    ##    #Branches #Junctions #End-pointvoxels #Junctionvoxels #Slabvoxels
    ## 1:        32         15               17              31         214
    ## 2:        29         13               17              33         165
    ## 3:        42         19               22              37         261
    ## 4:        19          8               11              14         145
    ## 5:        17          8               10              18         112
    ## 6:        26         11               16              26         184
    ##    AverageBranchLength #Triplepoints #Quadruplepoints MaximumBranchLength
    ## 1:               5.555            13                2              20.424
    ## 2:               5.103            12                0              12.618
    ## 3:               5.225            15                3              18.062
    ## 4:               6.340             6                1              18.203
    ## 5:               5.689             8                0              16.703
    ## 6:               6.095             8                3              23.083
    ##    LongestShortestPath SkelArea Ibranches(inferred) Intersectingradii
    ## 1:              71.090   88.137                   2                45
    ## 2:              61.586   72.326                   1                46
    ## 3:              72.768  107.648                   1                45
    ## 4:              67.452   57.188                   1                29
    ## 5:              41.867   47.096                   1                32
    ## 6:              72.809   76.026                   2                41
    ##    Suminters. Meaninters. Medianinters. Skewness(sampled) Kurtosis(sampled)
    ## 1:        182        4.04           4.0              0.58             -0.10
    ## 2:        157        3.41           3.5              0.45             -0.40
    ## 3:        235        5.22           5.0              0.27             -0.99
    ## 4:        109        3.76           4.0              0.07             -0.79
    ## 5:        103        3.22           3.0              0.36             -1.26
    ## 6:        156        3.80           4.0              0.67              1.02
    ##    Maxinters. Maxinters.radius Ramificationindex(sampled) Centroidradius
    ## 1:         10            12.73                          5          18.51
    ## 2:          8            10.39                          8          20.45
    ## 3:         12            13.87                         12          15.93
    ## 4:          7             7.63                          7          12.88
    ## 5:          7             8.67                          7          10.29
    ## 6:          8             4.98                          4          20.71
    ##    Centroidvalue Enclosingradius Criticalvalue Criticalradius Meanvalue
    ## 1:          5.31           29.55          7.27          13.45      4.10
    ## 2:          4.78           29.53          5.62          10.04      3.47
    ## 3:          5.43           30.11         10.24          13.29      5.31
    ## 4:          4.11           20.39          6.59           7.42      3.85
    ## 5:          2.84           20.85          5.96           9.34      3.28
    ## 6:          5.25           27.02          7.24           5.49      3.86
    ##    Ramificationindex(fit) Skewness(fit) Kurtosis(fit) Polyn.degree
    ## 1:                   3.64          0.16         -1.20            7
    ## 2:                   5.62          0.05         -1.24            8
    ## 3:                  10.24          0.13         -1.30            7
    ## 4:                   6.59          0.37         -0.53            8
    ## 5:                   5.96          0.06         -1.61            8
    ## 6:                   3.62          0.18          0.57            7
    ##    Regressioncoefficient(Semi-log) Regressionintercept(Semi-log)
    ## 1:                            0.18                         -2.21
    ## 2:                            0.18                         -2.41
    ## 3:                            0.18                         -2.17
    ## 4:                            0.20                         -2.25
    ## 5:                            0.24                         -2.08
    ## 6:                            0.20                         -2.12
    ##    Regressioncoefficient(Semi-log)[P10-P90]
    ## 1:                                     0.18
    ## 2:                                     0.19
    ## 3:                                     0.19
    ## 4:                                     0.22
    ## 5:                                     0.26
    ## 6:                                     0.17
    ##    Regressionintercept(Semi-log)[P10-P90] Regressioncoefficient(Log-log)
    ## 1:                                  -2.14                           2.40
    ## 2:                                  -2.24                           2.27
    ## 3:                                  -1.77                           2.30
    ## 4:                                  -1.98                           2.11
    ## 5:                                  -1.73                           2.11
    ## 6:                                  -2.49                           2.48
    ##    Regressionintercept(Log-log) Regressioncoefficient(Log-log)[P10-P90]
    ## 1:                         1.14                                    2.60
    ## 2:                         0.67                                    2.64
    ## 3:                         1.04                                    2.73
    ## 4:                         0.35                                    2.46
    ## 5:                         0.02                                    2.47
    ## 6:                         1.34                                    2.38
    ##    Regressionintercept(Log-log)[P10-P90] Density=ForegroundPixels/HullArea
    ## 1:                                  1.91                            0.3795
    ## 2:                                  1.87                            0.3459
    ## 3:                                  2.55                            0.4249
    ## 4:                                  1.33                            0.4986
    ## 5:                                  1.14                            0.5921
    ## 6:                                  1.23                            0.3927
    ##    SpanRatio(major/minoraxis) MaximumSpanAcrossHull      Area
    ## 1:                     1.6004              52.92001 1187.8284
    ## 2:                     1.5544              48.58863 1075.4708
    ## 3:                     1.0666              48.81999 1323.7340
    ## 4:                     1.0969              30.82741  597.7828
    ## 5:                     1.2966              31.79438  527.8116
    ## 6:                     1.2183              41.50550  915.6808
    ##    HullandCircularityPerimeter Circularity MaximumRadiusfromHull'sCentreofMass
    ## 1:                   133.77253      0.8341                            28.08517
    ## 2:                   126.20481      0.8485                            26.72066
    ## 3:                   138.84643      0.8629                            24.98820
    ## 4:                    91.88679      0.8897                            17.00032
    ## 5:                    89.44766      0.8290                            17.04643
    ## 6:                   114.11674      0.8836                            22.75578
    ##    Max/MinRadii CVforallRadii MeanRadius DiameterofBoundingCircle
    ## 1:       1.7271        0.1870   22.61704                 52.92001
    ## 2:       1.4612        0.1307   21.49822                 48.58863
    ## 3:       1.3792        0.0704   23.13260                 48.86239
    ## 4:       1.6197        0.1141   15.05384                 32.44642
    ## 5:       1.8084        0.1520   15.11956                 32.68393
    ## 6:       1.4597        0.1167   18.81288                 41.66134
    ##    MaximumRadiusfromCircle'sCentre Max/MinRadiifromCircle'sCentre
    ## 1:                        26.46001                         2.2254
    ## 2:                        24.29434                         1.6274
    ## 3:                        24.43117                         1.3086
    ## 4:                        16.22318                         1.4536
    ## 5:                        16.34196                         1.5931
    ## 6:                        20.83064                         1.5347
    ##    CVforallRadiifromCircle'sCentre MeanRadiusfromCircle'sCentre
    ## 1:                          0.2407                     22.12306
    ## 2:                          0.1598                     21.79675
    ## 3:                          0.0684                     23.13371
    ## 4:                          0.1048                     15.00332
    ## 5:                          0.1222                     15.10042
    ## 6:                          0.0961                     19.26609
    ##    FractalDimension Lacunarity CellNo BranchingDensity      InfInd
    ## 1:             1231       1039      1       0.07420011  0.02277333
    ## 2:              515       1698      2       0.06725055  1.50402880
    ## 3:             1480       1326      3       0.08132147 -1.01412498
    ## 4:             1401        806      4       0.09566685  0.73192524
    ## 5:             1818        761      5       0.08922881  0.06343808
    ## 6:             1342        222      6       0.08302675  0.43285031
